A Large-Scale Study of Polymorphism in R

### 2019-04-06 paper/main.tex
merging

\begin{abstract}
The R programming language is widely used in a variety of scientific domains
for tasks related to data science. The language was designed to favor an
interactive style of programming with minimal syntactic and conceptual
overhead. This design is well suited to support interactive data analysis,
but is not well suited to generating performant code or catching programming
errors.  In particular, R has no type annotations and all operations are
dynamically checked at runtime. The starting point for our work is the
question: \emph{what could a static type system for R look like?}  To answer
it, we study the polymorphism that is present in \RLOC lines of R
code spread among some \PACKAGES packages, written over a
period of over \YEARS years by thousands of programmers.  We perform a dynamic
analysis, leveraging tests and use-cases, to determine the level of
polymorphism that is present in the code. We do this for several potential
notions of types. Our results suggest that polymorphism is important in some
key parts of the system but that relatively simple type annotations could be
used to capture most of the interesting cases.

>>>>

\begin{abstract}
The R programming language is widely used in a variety of scientific domains
for tasks related to data science. The language was designed to favor an
interactive style of programming with minimal syntactic and conceptual
overhead. This design is well suited to support interactive data analysis,
but is not well suited to generating performant code or catching programming
errors.  In particular, R has no type annotations and all operations are
dynamically checked at runtime. The starting point for our work is the
question: \emph{what could a static type system for R look like?}  In order to answer
it, we study the polymorphism that is present in \RLOC lines of R
code spread among some \PACKAGES packages, written over a
period of over \YEARS years by thousands of programmers.  We perform a dynamic
analysis, leveraging tests and use-cases, to determine the level of
polymorphism that is present in the code. We do this for several potential
notions of types. Our results suggest that polymorphism is important in some
key parts of the system but that relatively simple type annotations could be
used to capture most of the interesting cases.
